---
title: "simpleRegress"
author: "Zoilo Franco"
date: "Spring 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 
    
## Model LENGTH ~ HEADLEN

```{r}
L1 <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(L1)
summary.lm(L1)
```
    
The plot shows a good linear relationship.    
The Adjusted R-Squared is 0.8434   
    
## Model LENGTH ~ HeadWTH
    
```{r}
L2 <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(L2)
summary.lm(L2)
```
    
The plot shows a good linear relationship.    
The Adjusted R-Squared is 0.5321

## Model LENGTH ~ CHEST

```{r}
L3 <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(L3)
summary.lm(L3)
```
    
The plot shows a good linear relationship.    
The Adjusted R-Squared is 0.7869   
    
## Model LENGTH ~ WEIGHT
    
```{r}
L4 <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(L4)
summary.lm(L4)
```
    
The plot shows a good linear relationship.    
The Adjusted R-Squared is 0.7423 
    
## Models from WORST to BEST
    
The models range from HEADWTH 0.5321, Weight 0.7423, CHEST.7869, to HEADLEN 0.8434 using Adjusted R-Squared. 
